MCC Models

### 2019-08-19 papers/full_paper/SI_appendix.Rmd
add analysis details to SI

---
title             : "Supplementary information:Pragmatic cue integration in adults’ and children’s inferences about novel word meanings"
shorttitle        : " "

@@

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F, include = T)
library(papaja)
library(tidyverse)
library(knitr)
library(ggthemes)
library(langcog)
library(rwebppl)
library(coda)
library(matrixStats)
library(ggpubr)
library(lme4)
library(broom)

@@

Based on euqation 2 we can now compute the likelihood of each utterance given an object type. Recall the two lexica: $\mathcal{L}_1 = \{dax: \text{pink-ish thing}, wug: \text{yellow-ish thing} \}$, $\mathcal{L}_2 = \{dax: \text{yellow-ish thing}, wug: \text{pink-ish thing}\}$, and the four utterances:  $u_1 = (\text{left}, dax)$, $u_2 = (\text{left}, wug)$, $u_3 = (\text{right}, dax)$, $u_4 = (\text{right}, wug)$.
We begin with the pink-ish type ($r_p$). If the speaker's lexicon were $\mathcal{L}_1$:

@@

To conclude the example, according to the parameter free pragmatics model, the proability that $L_1$ thinks that $S_1$ is reffering to a yellow-ish object when producing $u_3$ is 0.91. Based on this model we thus expected that, on average, participants would select the yellow-ish object in 91% of cases in a condition with the prior distribution specified above.

@@

The prior only model ignored the information about the intended referent that was expressed by the utterance and instead only focused on common ground manipulation. The only inforamtion available to $L_1$ is the prior distribution over referents. It is therefore defined as:

@@

# Prior strength manipualtions Experiment 4

apa_table(
  ex4_prior_models_table
  , caption = "Model output for prior strength experiments "
  , note = "Model structure in all cases: correct ~ condition + (1|id) + (1|agent)"
  , escape = T
)
```

Table \ref{tab:table_prior} shows the results of a generalized linear mixed model (GLMM) fit to the data from each manipulation. The results show that the parameter estimates for condition (i.e. difference between same speaker and different speaker condition) decreases in line with the hypothesized effect of the prior manipulation.

@@

>>>>

---
title             : "Supplementary information:Pragmatic cue integration in adults’ and children’s inferences about novel word meanings"
shorttitle        : "Supplementary information"

@@

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F, include = T)
library(papaja)
library(tidyverse)
library(knitr)
library(ggthemes)
library(langcog)
library(rwebppl)
library(coda)
library(matrixStats)
library(ggpubr)
library(lme4)
library(broom)
library(exactRankTests)

@@

Based on equation 2 we can now compute the likelihood of each utterance given an object type. Recall the two lexica: $\mathcal{L}_1 = \{dax: \text{pink-ish thing}, wug: \text{yellow-ish thing} \}$, $\mathcal{L}_2 = \{dax: \text{yellow-ish thing}, wug: \text{pink-ish thing}\}$, and the four utterances:  $u_1 = (\text{left}, dax)$, $u_2 = (\text{left}, wug)$, $u_3 = (\text{right}, dax)$, $u_4 = (\text{right}, wug)$.
We begin with the pink-ish type ($r_p$). If the speaker's lexicon were $\mathcal{L}_1$:

@@

To conclude the example, according to the parameter free pragmatics model, the probability that $L_1$ thinks that $S_1$ is referring to a yellow-ish object when producing $u_3$ is 0.91. Based on this model we thus expected that, on average, participants would select the yellow-ish object in 91% of cases in a condition with the prior distribution specified above.

@@

The prior only model ignored the information about the intended referent that was expressed by the utterance and instead only focused on common ground manipulation. The only information available to $L_1$ is the prior distribution over referents. It is therefore defined as:

@@

@@

# Supplemental results

The following sections present details on the analyses used to evaluate whether the different experimental manipulations produced different responses. For all generalized linear mixed models (GLMM) we used the maximally converging random effects structure.

## Experiment 1

```{r table_chance_Ex_1_2,results = "asis"}
chance_comp_ex_1_2 <- bind_rows(
    adult_ex1_data,
    adult_ex2_novelty_data %>% mutate(condition = ifelse(condition == "same_speaker", "test", "control")),
    adult_ex2_preference_data %>% mutate(condition = ifelse(condition == "same_speaker", "test", "control"))
  )%>%
    mutate(condition = factor(condition, levels = c("test","control")),
           experiment = factor(experiment, levels = c("adult_ex1","adult_ex2_preference","adult_ex2_novelty")),
           experiment = recode(experiment, 
                               adult_ex1 = "Experiment 1",
                               adult_ex2_preference = "Experiment 2A: Preference", 
                               adult_ex2_novelty = "Experiment 2B: Novelty"))%>%
  group_by(experiment,condition, id) %>%
  summarise(correct = mean(correct))%>%
  summarise(correct = list(correct)) %>%
  group_by(experiment,condition) %>%
  mutate(mean = mean(unlist(correct)),
         df= t.test(unlist(correct), mu = 0.5)$parameter,
         t_value = t.test(unlist(correct), mu = 0.5)$statistic,
         p_value = t.test(unlist(correct), mu = 0.5)$p.value) %>%
  select(experiment,condition,mean,df,t_value,p_value)%>%
  mutate(p_value = ifelse(p_value<.001,"< .001",as.character(paste("=",substr(round(p_value,3),2,5),sep=" "))),
         t_value = round(t_value,2))

apa_table(
  chance_comp_ex_1_2
  , caption = "Comparison to chance in each condition of Experiment 1, 2A and 2B"
  , note = "Proportion expected by chance = 0.5. Data aggregated within participant and conditon."
  , escape = T
)

We used one sample t-tests to test if participants selected the more informative object (the one unique to the table the animal pointed at) above chance (50%). Table \ref{tab:table_chance_Ex_1_2} shows the results.

```{r}
# model
lm_ex1 <- glmer(correct ~ condition +
              (1 |id), 
              data = adult_ex1_data, 
              family = binomial,control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))

#summary(lm_ex1)

ex1_r <- tidy(lm_ex1, effects = "fixed")%>%
  mutate(p.value = ifelse(p.value<.001,"< .001",as.character(paste("=",substr(round(p.value,3),2,5),sep=" "))),
         std.error = round(std.error,2),
         estimate = round(estimate,2))
```

To compare the two conditions, we fit the following GLMM: `correct ~ condition + (1 |id)`. Participants chose the more informative object more often in the test condition compared to the control condition ($\beta$ = `r ex1_r%>%filter(term == "conditiontest")%>%pull(estimate)`, se = `r ex1_r%>%filter(term == "conditiontest")%>%pull(std.error)`, *p* `r ex1_r%>%filter(term == "conditiontest")%>%pull(p.value)`).

## Experiment 2

```{r}
# model (maximally converging)
lm_ex2pref <- glmer(correct ~ condition +
              (1|id) + (1|agent), 
              data = adult_ex2_preference_data, 
              family = binomial,control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))

#summary(lm_ex2pref)

ex2pref_r <- tidy(lm_ex2pref, effects = "fixed")%>%
  mutate(p.value = ifelse(p.value<.001,"< .001",as.character(paste("=",substr(round(p.value,3),2,5),sep=" "))),
         std.error = round(std.error,2),
         estimate = round(estimate,2))

ex2tnov <- adult_ex2_novelty_data %>%
  group_by(condition, id) %>%
  summarise(correct = mean(correct)) %>%
  summarise(correct = list(correct)) %>%
  group_by(condition) %>%
  mutate(df= t.test(unlist(correct), mu = 0.5)$parameter,
         t_value = t.test(unlist(correct), mu = 0.5)$statistic,
         p_value = t.test(unlist(correct), mu = 0.5)$p.value) %>%
  select(condition,df,t_value,p_value)%>%
  mutate(p_value = ifelse(p_value<.001,"< .001",as.character(paste("=",substr(round(p_value,3),2,5),sep=" "))),
         t_value = round(t_value,2))

# model
lm_ex2nov <- glmer(correct ~ condition +
              (condition|id), 
              data = adult_ex2_novelty_data, 
              family = binomial,control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))

#summary(lm_ex2nov)

ex2nov_r <- tidy(lm_ex2nov, effects = "fixed")%>%
  mutate(p.value = ifelse(p.value<.001,"< .001",as.character(paste("=",substr(round(p.value,3),2,5),sep=" "))),
         std.error = round(std.error,2),
         estimate = round(estimate,2))
```

Table \ref{tab:table_chance_Ex_1_2} shows results for the comparison to chance in Experiment 2A and B. In Experiment 2A, we fit the following GLMM: `correct ~ condition + (1|id) + (1|agent)`. Here, `agent` refers to the animal making the request. Participants chose the object previously preferred by the speaker more often in the test (same speaker) condition compared to the control (different speaker) condition ($\beta$ = `r ex2pref_r%>%filter(term == "conditionsame_speaker")%>%pull(estimate)`, se = `r ex2pref_r%>%filter(term == "conditionsame_speaker")%>%pull(std.error)`, *p* `r ex2pref_r%>%filter(term == "conditionsame_speaker")%>%pull(p.value)`). For Experiment 2B, we fit the following GLMM: `correct ~ condition + (condition|id)`. Participants chose the novel object more often in the test (same speaker) condition compared to the control (different speaker) condition ($\beta$ = `r ex2nov_r%>%filter(term == "conditionsame_speaker")%>%pull(estimate)`, se = `r ex2nov_r%>%filter(term == "conditionsame_speaker")%>%pull(std.error)`, *p* `r ex2nov_r%>%filter(term == "conditionsame_speaker")%>%pull(p.value)`).

```{r ex3_model, results = "asis"}
# model 
lm_ex3 <- glmer(correct_inf ~ common_ground_manipulation*speaker*alignment + (alignment|id), 
              data = adult_ex3_data, family = binomial, 
              control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))

#summary(lm_ex3)

lm_ex3_r <- tidy(lm_ex3, effects = "fixed")%>%
  mutate(p.value = ifelse(p.value<.001,"< .001",as.character(paste("=",substr(round(p.value,3),2,5),sep=" "))),
         std.error = round(std.error,2),
         estimate = round(estimate,2))%>%
  as_tibble()%>%
  mutate(term = recode(term, 
                       `(Intercept)` = "Intercept",
                       common_ground_manipulationpreference = "common ground",
                       speakersame_speaker = "speaker",
                       alignmentincongruent = "alignment",
                       `common_ground_manipulationpreference:speakersame_speaker` = "common ground * speaker",
                       `common_ground_manipulationpreference:alignmentincongruent` = "common ground * alignment",
                       `speakersame_speaker:alignmentincongruent` = "speaker * alignment",
                       `common_ground_manipulationpreference:speakersame_speaker:alignmentincongruent` = "common ground * speaker * alignment"
         ))

apa_table(
  lm_ex3_r
  , caption = "GLMM output for Experiment 3"
  , note = "Reference levels: commmon ground - novelty, speaker - different speaker, alignment - congruent."
  , escape = T
)

To see if participants differentiated between conditions, we fit the following GLMM to the data: `correct_inf ~ common ground*speaker*alignment + (alignment|id)`. The dependent variable `correct_inf` captures whether participants chose the more informative object. Table \ref{tab:ex3_model} shows the model output.

## Prior strength manipualtions Experiment 4

apa_table(
  ex4_prior_models_table
  , caption = "GLMMs output for prior strength experiments"
  , note = "Model structure in all cases: correct ~ condition + (1|id) + (1|agent)"
  , escape = T
)
```

Table \ref{tab:table_prior} shows the results of a GLMM fit to the data from each manipulation. The results show that the parameter estimates for condition (i.e. difference between same speaker and different speaker condition) decreases in line with the hypothesized effect of the prior manipulation.

```{r ex4_models, results = "asis"}
lm_ex4_strong <- glmer(correct_inf ~ common_ground_manipulation*speaker*alignment + (speaker+alignment|id), 
              data = adult_ex4_data %>% filter(prior_manipulation == "strong"), family = binomial, 
              control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))

#summary(lm_ex4_strong)

lm_ex4_r_strong <- tidy(lm_ex4_strong, effects = "fixed")%>%
  mutate(p.value = ifelse(p.value<.001,"< .001",as.character(paste("=",substr(round(p.value,3),2,5),sep=" "))),
         std.error = round(std.error,2),
         estimate = round(estimate,2))%>%
  as_tibble()%>%
  mutate(term = recode(term, 
                       `(Intercept)` = "Intercept",
                       common_ground_manipulationpreference = "common ground",
                       speakersame_speaker = "speaker",
                       alignmentincongruent = "alignment",
                       `common_ground_manipulationpreference:speakersame_speaker` = "common ground * speaker",
                       `common_ground_manipulationpreference:alignmentincongruent` = "common ground * alignment",
                       `speakersame_speaker:alignmentincongruent` = "speaker * alignment",
                       `common_ground_manipulationpreference:speakersame_speaker:alignmentincongruent` = "common ground * speaker * alignment"
         ))%>%
  mutate(`prior strength` = "strong")%>%
  select(`prior strength`, term, estimate, std.error, statistic,p.value)

lm_ex4_medium <- glmer(correct_inf ~ common_ground_manipulation*speaker*alignment + (alignment|id), 
              data = adult_ex4_data %>% filter(prior_manipulation == "medium"), family = binomial, 
              control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))

#summary(lm_ex4_medium)

lm_ex4_r_medium <- tidy(lm_ex4_medium, effects = "fixed")%>%
  mutate(p.value = ifelse(p.value<.001,"< .001",as.character(paste("=",substr(round(p.value,3),2,5),sep=" "))),
         std.error = round(std.error,2),
         estimate = round(estimate,2))%>%
  as_tibble()%>%
  mutate(term = recode(term, 
                       `(Intercept)` = "Intercept",
                       common_ground_manipulationpreference = "common ground",
                       speakersame_speaker = "speaker",
                       alignmentincongruent = "alignment",
                       `common_ground_manipulationpreference:speakersame_speaker` = "common ground * speaker",
                       `common_ground_manipulationpreference:alignmentincongruent` = "common ground * alignment",
                       `speakersame_speaker:alignmentincongruent` = "speaker * alignment",
                       `common_ground_manipulationpreference:speakersame_speaker:alignmentincongruent` = "common ground * speaker * alignment"
         ))%>%
  mutate(`prior strength` = "medium")%>%
  select(`prior strength`, term, estimate, std.error, statistic,p.value)

lm_ex4_weak <- glmer(correct_inf ~ speaker*alignment + (speaker+alignment|id), 
              data = adult_ex4_data %>% filter(prior_manipulation == "weak"), family = binomial, 
              control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))

#summary(lm_ex4_weak)

lm_ex4_r_weak <- tidy(lm_ex4_weak, effects = "fixed")%>%
  mutate(p.value = ifelse(p.value<.001,"< .001",as.character(paste("=",substr(round(p.value,3),2,5),sep=" "))),
         std.error = round(std.error,2),
         estimate = round(estimate,2))%>%
  as_tibble()%>%
  mutate(term = recode(term, 
                       `(Intercept)` = "Intercept",
                       common_ground_manipulationpreference = "common ground",
                       speakersame_speaker = "speaker",
                       alignmentincongruent = "alignment",
                       `common_ground_manipulationpreference:speakersame_speaker` = "common ground * speaker",
                       `common_ground_manipulationpreference:alignmentincongruent` = "common ground * alignment",
                       `speakersame_speaker:alignmentincongruent` = "speaker * alignment",
                       `common_ground_manipulationpreference:speakersame_speaker:alignmentincongruent` = "common ground * speaker * alignment"
         ))%>%
  mutate(`prior strength` = "weak")%>%
  select(`prior strength`, term, estimate, std.error, statistic,p.value)

t_ex4_r <- bind_rows(
  lm_ex4_r_strong,
  lm_ex4_r_medium,
  lm_ex4_r_weak
)

apa_table(
  t_ex4_r
  , caption = "GLMMs output for Experiment 4"
  , note = "Reference levels: commmon ground - novelty, speaker - different speaker, alignment - congruent."
  , escape = T
)

To assess how participants differentiated between combinations of manipulations, we fit a separate GLMM to the data of each level of prior strength manipulation. Please note that we did not find a weak manipulation for preference. Model outputs are shown in Table \ref{tab:ex4_models}. Model structure for each level of prior strength manipulation was as follows:

* Strong:`correct_inf ~ common ground*speaker*alignment + (speaker+alignment|id)`
* Medium:`correct_inf ~ common ground*speaker*alignment + (alignment|id)`
* Weak:`correct_inf ~ correct_inf ~ speaker*alignment + (speaker+alignment|id)`

## Experiment 5

```{r child_chance_comp, results = "asis"}
#experiment 1
child_ex1t <- child_ex1_data %>%
  mutate(condition = "test",
         experiment = "Experiment 5")%>%
  group_by(experiment,condition,age_bin, id) %>%
  summarise(correct = mean(correct)) %>%
  summarise(correct = list(correct)) %>%
  group_by(age_bin) %>%
  mutate(mean = mean(unlist(correct)),
         df= t.test(unlist(correct), mu = 0.5)$parameter,
         t_value = t.test(unlist(correct), mu = 0.5)$statistic,
         p_value = t.test(unlist(correct), mu = 0.5)$p.value) %>%
  select(experiment,condition, mean, age_bin,df,t_value,p_value)%>%
  mutate(p_value = ifelse(p_value<.001,"< .001",as.character(paste("=",substr(round(p_value,3),2,5),sep=" "))),
         t_value = round(t_value,2))

child_ex1_lm_data <- child_ex1_data%>%
  mutate(age_num = scale(age_num, center = TRUE, scale = TRUE))

# experiment 2
child_ex2t <- child_ex2_data %>%
  mutate(condition = ifelse(condition == "same_speaker", "test", "control"),
         experiment = "Experiment 6")%>%
  group_by(experiment,condition,age_bin, id) %>%
  summarise(correct = mean(correct)) %>%
  summarise(correct = list(correct)) %>%
  group_by(condition,age_bin) %>%
  mutate(mean = mean(unlist(correct)),
         df= t.test(unlist(correct), mu = 0.5)$parameter,
         t_value = t.test(unlist(correct), mu = 0.5)$statistic,
         p_value = t.test(unlist(correct), mu = 0.5)$p.value) %>%
  select(experiment, condition, mean, age_bin,df,t_value,p_value)%>%
  mutate(p_value = ifelse(p_value<.001,"< .001",as.character(paste("=",substr(round(p_value,3),2,5),sep=" "))),
         t_value = round(t_value,2))

child_chance <- 
  bind_rows(
    child_ex1t,
    child_ex2t
  )

apa_table(
  child_chance
  , caption = "Comparison to chance in each condition of Experiment 5 and 6"
  , note = "Proportion expected by chance = 0.5. Data aggregated within participant and conditon."
  , escape = T
)

Data for children was binned by year for the comparison to chance. As for adults, we used one sample t-tests for this analysis. Table \ref{tab:child_chance_comp} shows the results for Experiment 5.

```{r}
child_ex1_lm_data <- child_ex1_data%>%
  mutate(age_num = scale(age_num, center = TRUE, scale = TRUE))

# GLMM
child_ex1_lm <- glmer(correct ~ age_num + (1 | id), 
              data = child_ex1_lm_data, family = binomial, 
              control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))

child_lm_ex1_r <- tidy(child_ex1_lm, effects = "fixed")%>%
  mutate(p.value = ifelse(p.value<.001,"< .001",as.character(paste("=",substr(round(p.value,3),2,5),sep=" "))),
         std.error = round(std.error,2),
         estimate = round(estimate,2))

We fit a GLMM with the following structure to the trial by trial data with age as a continuous variable: `correct ~ age_num + (1|id)`. Results show that children became more likely to select the more informative object with age ($\beta$ = `r child_lm_ex1_r%>%filter(term == "age_num")%>%pull(estimate)`, se = `r child_lm_ex1_r%>%filter(term == "age_num")%>%pull(std.error)`, *p* `r child_lm_ex1_r%>%filter(term == "age_num")%>%pull(p.value)`.

## Experiment 6

```{r}
child_ex2_lm_data <- child_ex2_data%>%
  mutate(age_num = scale(age_num, center = TRUE, scale = TRUE))

child_ex2_lm <- glmer(correct ~ age_num*condition + (condition | id) + (1 | agent), 
              data = child_ex2_lm_data, family = binomial, 
              control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))

child_lm_ex2_r <- tidy(child_ex2_lm, effects = "fixed")%>%
  mutate(p.value = ifelse(p.value<.001,"< .001",as.character(paste("=",substr(round(p.value,3),2,5),sep=" "))),
         std.error = round(std.error,2),
         estimate = round(estimate,2))
```

Comparison to chance within each condition of Experiment 6 is given in Table \ref{tab:child_chance_comp}. We fit a GLMM with the following structure to the data: `correct ~ age_num*condition + (condition|id) + (1|agent)`. The model output shows an effect of condition (same or different speaker) ($\beta$ = `r child_lm_ex2_r%>%filter(term == "conditionsame_speaker")%>%pull(estimate)`, se = `r child_lm_ex2_r%>%filter(term == "conditionsame_speaker")%>%pull(std.error)`, *p* `r child_lm_ex2_r%>%filter(term == "conditionsame_speaker")%>%pull(p.value)`) but no effect of age ($\beta$ = `r child_lm_ex2_r%>%filter(term == "age_num")%>%pull(estimate)`, se = `r child_lm_ex2_r%>%filter(term == "age_num")%>%pull(std.error)`, *p* `r child_lm_ex2_r%>%filter(term == "age_num")%>%pull(p.value)`) or interaction between speaker identity and age ($\beta$ = `r child_lm_ex2_r%>%filter(term == "age_num:conditionsame_speaker")%>%pull(estimate)`, se = `r child_lm_ex2_r%>%filter(term == "age_num:conditionsame_speaker")%>%pull(std.error)`, *p* `r child_lm_ex2_r%>%filter(term == "age_num:conditionsame_speaker")%>%pull(p.value)`).

```{r child_models, results = "asis"}
child_ex3_lm_data <- child_ex3_data%>%
  mutate(age_num = scale(age_num, center = TRUE, scale = TRUE))

child_ex3_lm <- glmer(correct_inf ~ age_num*speaker*alignment 
      + (speaker+alignment | id), 
      family = "binomial",
      data = child_ex3_lm_data,
      control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))

child_lm_ex3_r <- tidy(child_ex3_lm, effects = "fixed")%>%
  mutate(p.value = ifelse(p.value<.001,"< .001",as.character(paste("=",substr(round(p.value,3),2,5),sep=" "))),
         std.error = round(std.error,2),
         estimate = round(estimate,2))%>%
    mutate(term = recode(term, 
                       `(Intercept)` = "Intercept",
                       age_num = "age",
                       speakersame_speaker = "speaker",
                       alignmentincongruent = "alignment",
                       `age_num:speakersame_speaker` = "age * speaker",
                       `age_num:alignmentincongruent` = "age * alignment",
                       `speakersame_speaker:alignmentincongruent` = "speaker * alignment",
                       `age_num:speakersame_speaker:alignmentincongruent` = "age * speaker * alignment"
         ))

apa_table(
  child_lm_ex3_r
  , caption = "GLMM output for Experiment 7"
  , note = "Reference level: speaker - different speaker, alignment - congruent."
  , escape = T
)
```

To see if children differentiated between the different combinations of manipulations in Experiment 7, we fit the following GLMM to the data: `correct_inf ~ age_num*speaker*alignment+ (speaker+alignment|id)`. Model output is shown in Table \ref{tab:child_models}.
